---
title: "CBASS 2020: Thermal tolerance of *A. cervicornis* across 6 Florida coral nurseries"
author: Ross Cunning
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)

library(readxl)
library(janitor)
library(parzer)
library(broom)
library(lsmeans)
library(drc)
library(tidyverse)
library(ggthemes)
library(multcompView)
library(ggpubr)
library(mcr)
```

This document analyzes thermal tolerance metrics (ED50) derived from each coral genotype measured at each nursery on the R/V *Coral Reef II* in the summer and fall of 2020. ED50 values are analyzed for differences among nurseries, and effects of source colony location and temperature regime. For a subset of corals, ED50 values are compared to values obtained from the same genets at the same nurseries in spring 2020, in a separate, independent set of CBASS assays.

```{r import_data}
# Import cleaned datasets from each nursery
clean_data_files <- list.files(path = "data/processed", 
                               pattern = "*_clean.csv", 
                               full.names = TRUE)
clean_data <- clean_data_files %>% map_dfr(read_csv)

# Import genotype metadata
gmd <- read_csv("data/genotype_metadata_raw.csv") %>%
  mutate(nursery = tolower(nursery),
         across(everything(), str_squish),
         across(ends_with("lat"), ~signif(parse_lat(.), 6)),             # parse latitude
         across(ends_with("lon"), ~signif(parse_lon(.), 6))) %>%        # parse longitude
  mutate(ugeno = gsub(" ", "", ugeno)) %>%
  select(nursery, name = geno, source_lat, source_lon, MMM_5km)

# Import names key
gkey <- read_csv("data/genotype_name_key.csv") %>%
  mutate(alt0 = geno) %>%
  pivot_longer(starts_with("alt"), names_to = "alt", values_to = "name") %>%
  drop_na() %>%
  select(geno, name)

# Combine cleaned data with genotype metadata
df.all <- clean_data %>%
  rename(name = geno) %>%
  left_join(gkey) %>%
  select(nursery, geno, name, max_temp, fvfm, date) %>%
  left_join(gmd) %>%
  # Order nurseries from south to north
  mutate(nursery = factor(nursery, levels = c("mote", "fwc", "rrt", "crf", "um", "nsu"))) %>%
  select(nursery, name, geno, date, max_temp, fvfm, source_lat, source_lon, MMM_5km)

# Define dose-response curves using same constraints
ll3 <- function(data) {
  drm(fvfm ~ max_temp, data = data, 
      fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 40),
      lowerl = c(10, 0.55, 30))}

# Fit model to each coral, get parameters, fitted values, and residuals
mods <- df.all %>%
  nest(data = c(max_temp, fvfm)) %>%
  # Fit the model to each coral
  mutate(ll3 = map(data, ll3)) %>%
  # Get model parameters and fitted values/residuals
  mutate(pars = map(ll3, tidy),
         pred = map2(ll3, data, ~augment(.x, drop_na(.y, fvfm))))

# Extract parameter values from model fits
res.all <- mods %>%
  unnest(pars) %>%
  pivot_wider(names_from = term, 
              values_from = c(estimate, std.error, statistic, p.value)) %>%
  clean_names() %>%
  select(!where(is.list), -curve)

# Filter to analyze only CBASS runs from Coral Reef II in August and October
res <- res.all %>%
  filter(date %in% c("2020-08", "2020-10"))
```

# How many corals and unique genotypes were tested?
```{r}
res %>%
  summarize(ncoral = n_distinct(nursery, name),
            ngeno = n_distinct(geno)) %>%
  knitr::kable()
```


```{r test_geno_effect, eval = FALSE, include = FALSE}
# Do different genotypes have different thermal tolerance?


# Fit single curve without genotype as a factor
mod_nogeno <- drm(fvfm ~ max_temp, data = df, 
      fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 40),
      lowerl = c(10, 0.55, 30))
# Fit curves separately for each genotype
mod_geno <- drm(fvfm ~ max_temp, curveid = geno, data = df,
                fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 40),
      lowerl = c(10, 0.55, 30))
# Compare model fit with vs. without genotype as factor
anova(mod_nogeno, mod_geno)

## Genotype is highly significant
```

# What is the range in thermal tolerance across all genotypes?

```{r ed50_histograms, fig.width = 11}
# Basic histogram
# ggplot(res, aes(x = estimate_ed50)) + 
#   geom_histogram(binwidth = 0.1) +
#   labs(x = "ED50 Tolerance (°C)", y = "Number of genotypes",
#        title = "Histogram of ED50 values for all corals") +
#   theme_few()

# Histogram colored by nursery with density plot
hist1 <- res %>%
  mutate(nursery = factor(nursery, levels = c("fwc", "um", "mote", "crf", "nsu", "rrt"))) %>%
  ggplot(aes(x = estimate_ed50)) + 
  geom_histogram(aes(fill = nursery), binwidth = 0.1, alpha = 0.4) +
  geom_density(aes(y = ..count.. * 0.1)) +
  theme_few() +
  theme(legend.position = c(0.15, 0.65), legend.key.size = unit(0.25, "cm")) +
  labs(x = "ED50 (°C)", y = "Count", title = expression(V[P]))

# Remove nursery effect
mod <- lm(estimate_ed50 ~ nursery, data = res)
res.adj <- augment(mod, data = res) %>%
  mutate(ed50_adj = mean(res$estimate_ed50) + .resid)

# Get mean and sd from adjusted distribution
meanadj <- mean(res.adj$ed50_adj)
sdadj <- sd(res.adj$ed50_adj)

hist2 <- ggplot(res.adj, aes(x = ed50_adj)) + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.1, alpha = 0.4) +
  stat_function(fun = ~ dnorm(.x, meanadj, sdadj) * 0.1) +
  geom_vline(aes(xintercept = meanadj)) +
  geom_segment(aes(x = meanadj, xend = meanadj + sdadj, 
                   y = dnorm(meanadj + sdadj, meanadj, sdadj) * 0.1,
                   yend = dnorm(meanadj + sdadj, meanadj, sdadj) * 0.1),
               arrow = arrow(length = unit(0.2,"cm"), ends = "both"), lwd = 0.1) +
  xlim(qnorm(c(0.00001, 0.99999), meanadj, sdadj)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  annotate("text", x = meanadj - 0.1, y = 0.125, adj = 1,
           label = paste0("mean = ", round(meanadj, 2))) +
  annotate("text", x = meanadj + sdadj + 0.1, 
           y = dnorm(meanadj + sdadj, meanadj, sdadj) * 0.1, 
           adj = 0, label = paste0("s.d. = ", round(sdadj, 2))) +
  theme_few() +
  labs(x = "ED50 (°C) adjusted", y = "Percentage", title = expression(V[P] - V[E]))


# Combine histograms into multi-panel figure
fig3 <- cowplot::plot_grid(
  hist1 + xlim(qnorm(c(0.00001, 0.99999), meanadj, sdadj)), 
  hist2, labels = "AUTO")
fig3

# Summarize original distribution
tidy(summary(res$estimate_ed50)) %>%
  knitr::kable(caption = "Summary of unadjusted ED50 values")

# Is adjusted distribution normal?
tidy(shapiro.test(res.adj$ed50_adj)) %>% 
  knitr::kable(caption = "Normality test for adjusted ED50 values")

ggsave("output/fig3.png", width = 172, height = 75, units = "mm")
```

After adjusting for variation due to nursery location (i.e., environment), ED50 values are approximately normally distributed, with a mean of 35.98°C and standard deviation of 0.38°C. This is the variance due to genetic effects (and measurement error).

# How does thermal tolerance vary across nurseries?

```{r nursery_effects, fig.width = 5}
# Test for equal variance across nurseries
car::leveneTest(estimate_ed50 ~ nursery, data = res) %>%
  tidy() %>%
  knitr::kable(caption = "Levene's test for homogeneity of variance")
## Unequal variance among nurseries -- use Welch ANOVA
welch <- oneway.test(estimate_ed50 ~ nursery, data = res)
tidy(welch) %>% knitr::kable(caption = "Welch ANOVA table")
# Variance explained by nursery
omega.squared <- function(WelchF, df, N) df * (WelchF - 1) / (df * (WelchF - 1) + N)
tibble(omega.squared = omega.squared(
  WelchF = welch$statistic[[1]], df = welch$parameter[[1]], N = nrow(res))) %>%
  knitr::kable(caption = "Variance explained by nursery")

# Get mean and sd by nursery for plotting
nurs.stats <- res %>%
  group_by(nursery) %>%
  summarise(mean = mean(estimate_ed50), sd = sd(estimate_ed50)) %>%
  ungroup()

# Median pairwise difference between nurseries
mpwdiff <- median(dist(nurs.stats$mean))
mpwdiff %>%
  knitr::kable(caption = "Median pairwise difference between nurseries (°C)")

# Use pairwise wilcox to test for differences in ED50 among nurseries
test <- pairwise.wilcox.test(res$estimate_ed50, res$nursery, p.adjust.method = "bonf")
pvals <- tibble(reshape2::melt(test$p.value)) %>% drop_na() %>%
  mutate(Var1 = factor(Var1, levels = levels(fct_reorder(nurs.stats$nursery, nurs.stats$mean)))) %>%
  arrange(Var1)
stats <- setNames(pull(pvals[,3]), paste(pvals$Var1, pvals$Var2, sep = "-")) %>%
  multcompLetters(threshold = 0.01)
lett <- enframe(stats$monospacedLetters, name = "nursery", value = "group")
nurs.stats <- left_join(nurs.stats, lett)


# Raincloud plot
source("https://raw.githubusercontent.com/datavizpyr/data/master/half_flat_violinplot.R")
nurs.fig <- ggplot(res, aes(x = nursery, y = estimate_ed50, fill = nursery, group = nursery)) +
  geom_hline(aes(yintercept = mean(estimate_ed50)), lty = 2, col = "gray") +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   adjust = 1, width = 2, alpha = 0.7, lwd = 0) +
  geom_boxplot(position = position_nudge(x = 0.2, y = 0), width = 0.1,
               outlier.shape = NA, alpha = 1, lwd = 0.25) +
  geom_point(aes(color = nursery), alpha = 0.7,
             position = position_jitter(width = 0.075)) +
  geom_point(data = nurs.stats, aes(x = nursery, y = mean), inherit.aes = FALSE,
             position = position_nudge(x = 0, y = 0), size = 2, pch = 5) +
  geom_errorbar(data = nurs.stats, inherit.aes = FALSE, width = 0, lwd = 0.25,
                aes(x = nursery, y = mean, ymin = mean - sd, ymax = mean + sd),
                position = position_nudge(x = 0, y = 0)) +
  geom_text(data = nurs.stats, aes(y = 37.5, label = group), size = 2) +
  coord_flip() +
  scale_size(range = c(0.5, 3)) +
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  theme_few() +
  theme(legend.position = "none") +
  labs(y = "ED50 (°C)", x = "Nursery")
nurs.fig

ggsave("output/fig4.png", width = 120, height = 100, units = "mm")
```

There is variation among nurseries in average ED50 values. This could be due to different environmental conditions, and/or different set of genotypes present, and/or confounding experimental effects. However, most of the variation in thermal tolerance is found *within* nurseries rather than between them. This means that each nursery tends to have genets with a wide range of thermal tolerance phenotypes.


# How does thermal tolerance vary for the same genotype grown at different nurseries?
```{r, include = T, eval = T}
# Get ED50s (adjusted for nursery effect) for genotypes present at multiple nurseries
multi.adj <- res.adj %>%
  # Filter genotypes present at more than one nursery
  group_by(geno) %>%
  filter(n_distinct(nursery) > 1) %>%
  # If genotype run more than once at a nursery, use mean ed50 within nursery
  group_by(geno, nursery) %>%
  summarize(ed50_adj = mean(ed50_adj)) %>%
  ungroup()
# Reorder genotypes based on mean ed50 across nurseries
multi.adj <- multi.adj %>%
  group_by(geno) %>%
  summarize(geno_mean_ed50 = mean(ed50_adj)) %>%
  right_join(multi.adj) %>%
  mutate(geno = fct_reorder(geno, geno_mean_ed50))

# multi.adj %>%
#   count(geno, name = "n_nurs") %>%
#   count(n_nurs, name = "n_geno")

# multi <- res %>%
#   group_by(geno) %>%
#   filter(n_distinct(nursery) > 1) %>%
#   # If genotype run more than once at a nursery, use mean ed50 within nursery
#   group_by(geno, nursery) %>%
#   summarize(estimate_ed50 = mean(estimate_ed50)) %>%
#   ungroup()
# 
# # Adjust for nursery differences
# multi.adj <- multi %>%
#   group_by(nursery) %>%
#   summarise(nurs_mean = mean(estimate_ed50)) %>%
#   ungroup() %>%
#   right_join(multi) %>%
#   mutate(.resid = estimate_ed50 - nurs_mean,
#          ed50_adj = mean(estimate_ed50) + .resid)
# multi.adj <- multi.adj %>%
#   # Reorder genotypes based on mean ed50 across nurseries
#   group_by(geno) %>%
#   summarize(geno_mean_ed50 = mean(ed50_adj)) %>%
#   right_join(multi.adj) %>%
#   mutate(geno = fct_reorder(geno, geno_mean_ed50))
  

# Plot ED50adj by genotype
ggplot(multi.adj, aes(x = geno, y = ed50_adj, color = nursery, shape = nursery, group = geno)) + 
  geom_point(size = 1.5, alpha = 0.75) + 
  geom_point(aes(y = geno_mean_ed50), pch = 1, color = "black", stroke = 0.25, size = 2)+
  stat_summary(fun.data = mean_se, geom = "errorbar", color = "black", lwd = 0.25, width = 0.25) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 8),
        legend.position = c(0.1, 0.8)) +
  labs(x = "Genotype", y = "Adjusted ED50 (°C)")

tidy(anova(lm(ed50_adj ~ geno, data = multi.adj))) %>%
  knitr::kable(caption = "ANOVA: Effect of genotype on ED50")

library(GGally)
cors <- multi.adj %>%
  select(nursery, geno, ed50_adj) %>%
  pivot_wider(names_from = nursery, values_from = ed50_adj, values_fn = mean) %>%
  select(-geno)

ggpairs(cors, title = "Correlation of ED50 values for same genotype at multiple nurseries")

corfig <- ggpairs(cors)
ggsave(plot = corfig, filename = "output/figsss.png", width = 190, height = 190, units = "mm")

# library(RVenn)
# overlap <- res.adj %>%
#   distinct(nursery, geno) %>%
#   {split(.$geno, .$nursery)} %>%
#   Venn(.) %>%
#   overlap_pairs(.) %>%
#   map(length) %>%
#   unlist()
# overlap
# hist(overlap)
# mean(overlap)
# median(overlap[overlap > 0])

#ggsave("output/fig5.png", width = 90, height = 75, units = "mm")

# # Randomize ed50 values
# multi.adj.r <- multi.adj %>%
#   select(-geno_mean_ed50) %>%
#   mutate(ed50_adj = sample(ed50_adj))
# multi.adj.r %>%
#   # Reorder genotypes based on mean ed50 across nurseries
#   group_by(geno) %>%
#   summarize(geno_mean_ed50 = mean(ed50_adj)) %>%
#   right_join(multi.adj.r) %>%
#   ungroup() %>%
#   mutate(geno = fct_reorder(geno, geno_mean_ed50)) %>%
#   ggplot(aes(x = geno, y = ed50_adj, color = nursery, shape = nursery, group = geno)) + 
#   geom_point(size = 1.5, alpha = 0.75) + 
#   geom_point(aes(y = geno_mean_ed50), pch = 1, color = "black", stroke = 0.25, size = 2)+
#   stat_summary(fun.data = mean_se, geom = "errorbar", color = "black", lwd = 0.25, width = 0.25) +
#   theme_few() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1),
#         axis.text = element_text(size = 8),
#         legend.position = c(0.1, 0.8)) +
#   labs(x = "Genotype", y = "Adjusted ED50 (°C)")
#   
# # Test for equal variance across genotypes
# car::leveneTest(ed50_adj ~ geno, data = multi.adj.r)    # p = 0.019 => cannot use classic ANOVA
# 
# # Test for differences among genotypes with Welch's ANOVA
# (welch <- oneway.test(ed50_adj ~ geno, data = multi.adj.r, var.equal = F))
# ## Significant effect of genotype. Interpretation: ED50 value has a strong genetic component, and thermal tolerance phenotypes for a genet are reproducible even when genet moved to new environment.
# 
# # Estimate variance explained by genotype
# omega.squared <- function(WelchF, df, N) df * (WelchF - 1) / (df * (WelchF - 1) + N)
# omega.squared(WelchF = welch$statistic[[1]], df = welch$parameter[[1]], N = nrow(multi.adj))
# ## Genotype explains 49% of variance in ED50 measured across nurseries. Residual variance may be due to genotype by environment interactions (e.g., differential acclimatization of genotypes to different nursery environments), or CBASS run variability & measurement error.
# anova(lm(ed50_adj ~ geno, data = multi.adj.r))
# 
# 
# 
# 
# # Use unadjusted data?
# # Get ED50s (adjusted for nursery effect) for genotypes present at multiple nurseries
# multi <- res %>%
#   # Filter genotypes present at more than one nursery
#   group_by(geno) %>%
#   filter(n_distinct(nursery) > 1) %>%
#   # If genotype run more than once at a nursery, use mean ed50 within nursery
#   group_by(geno, nursery) %>%
#   summarize(ed50 = mean(estimate_ed50)) %>%
#   ungroup()
# # Reorder genotypes based on mean ed50 across nurseries
# multi <- multi %>%
#   group_by(geno) %>%
#   summarize(geno_mean_ed50 = mean(ed50)) %>%
#   right_join(multi) %>%
#   mutate(geno = fct_reorder(geno, geno_mean_ed50))
# 
# mod <- lm(ed50 ~ nursery + geno, data =  multi)
# anova(mod)
# car::Anova(mod, type = 3, white.adjust = TRUE)
# oneway.test(ed50 ~ geno, data = mutate(multi, ed50 = sample(ed50)), var.equal = F)
# 
# # These tests are really unstable with randomized data, don't think results are reliable. Sample sizes for each geno are so small.
```

Genotype is not a significant predictor of ED50 among coral genotypes tested at multiple nurseries. This could indicate that: 1) ED50 is plastic and genotype by environment interactions result in genotypes having different ED50 values at different nurseries, and/or 2) small sample sizes for each genotype limit statistical power to detect differences among genotypes. The 30 genotypes in this subset were present at only two (n=17) or three nurseries (n=13), so the low replication for each genotype is definitely a limitation. This study was not really designed to be able to answer this question, but the genotype swap experiment should hopefully be able to!


```{r, include = FALSE, eval = FALSE}
# Is thermal tolerance related to symbiont density?

sh <- read_csv("data/processed/acer_SH_data.csv") %>%
  clean_names() %>%
  filter(species == "Acer") %>%
  mutate(nursery = tolower(nursery)) %>%
  rename(name = genotype) %>%
  left_join(res.adj, sh, by = c("nursery", "name")) %>%
  mutate(a_acer = replace_na(a_acer, 0),
         d_acer = replace_na(d_acer, 0),
         totSH = a_acer + d_acer,
         logSH = log10(totSH)) %>%
  filter(nursery != "rrt")

ggplot(sh, aes(x = logSH, y = estimate_ed50)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

ggplot(sh, aes(x = logSH, y = estimate_ed50, color = nursery)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

mod <- lm(estimate_ed50 ~ logSH * nursery, data = sh)
anova(mod)
plot(emtrends(mod, specs = "nursery", var = "logSH"))
```


# Is thermal tolerance related to source colony latitude and longitude?
```{r}
mod <- lm(ed50_adj ~ source_lat + source_lon, data = res.adj)
anova(mod) %>%
  tidy() %>%
  knitr::kable()

lat.plot <- ggplot(res.adj, aes(x = source_lat, y = ed50_adj)) +
  geom_point(size = 0.75) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_few() +
  theme(axis.text = element_text(size = 7)) +
  labs(x = "Latitude", y = expression(paste(ED50[adj], " (°C)")))

lon.plot <- ggplot(res.adj, aes(x = source_lon, y = ed50_adj)) +
  geom_point(size = 0.75) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_few() +
  theme(axis.text = element_text(size = 7)) +
  labs(x = "Longitude", y = NULL)

cowplot::plot_grid(lat.plot, lon.plot, ncol = 2)
```

# Is thermal tolerance related to source colony temperature regime?

To test this hypothesis, we calculate maximum monthly mean (MMM) temperatures corresponding to each source colony location, using satellite SST products with both 9km and 5km resolution. 

```{r, eval = TRUE}
# Import 9km SST data
sst9 <- tibble(
  filename = list.files("data/sst_9km", full.names = TRUE),
  data = map(filename, read_csv, col_names = c("yyyymm", "temp")),
  coords = map(filename, ~ str_split_fixed(basename(.), "_", 4)),
  lat = map_dbl(coords, ~ as.numeric(nth(., 3))),
  lon = map_dbl(coords, ~ parse_number(nth(., 4)))
) %>%
  select(lat, lon, data)

sst9 <- sst9 %>%
  mutate(data = map(data, ~ mutate(., year = str_sub(yyyymm, 1, 4), month = str_sub(yyyymm, 5, 6)))) %>%
  mutate(monthly_means = map(data, ~ group_by(., month) %>% summarize(mean = mean(temp))),
         mmm_9km = map_dbl(monthly_means, ~ max(.$mean)),
         minmm = map_dbl(monthly_means, ~ min(.$mean)),
         totvar = map_dbl(data, ~ var(.$temp)),
         sumvar = map_dbl(data, ~ filter(.,  month %in% c("07", "08", "09")) %>% 
                            summarize(var = var(temp)) %>% pull(var)))

# Join with results and sample metadata
dft <- sst9 %>%
  distinct(source_lat = lat, source_lon = lon, mmm_9km, minmm, totvar, sumvar) %>%
  right_join(res.adj) %>%
  mutate(mmm_5km = as.numeric(mmm_5km))

# Test effect of MMM from 9km data
mod9km <- lm(ed50_adj ~ mmm_9km, data = dft)
anova(mod9km) %>%
  tidy() %>%
  knitr::kable()

# Test effect of MMM from 5km data
mod5km <- lm(ed50_adj ~ mmm_5km, data = dft)
anova(mod5km) %>%
  tidy() %>%
  knitr::kable()

# Plots
mmm9.plot <- ggplot(dft, aes(x = mmm_9km, y = ed50_adj)) +
  geom_point(size = 0.75) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_few() +
  theme(axis.text = element_text(size = 7)) +
  scale_x_continuous(labels = function(x) str_trim(x)) +
  labs(x = "MMM (9 km)", y = NULL)

mmm5.plot <- ggplot(dft, aes(x = mmm_5km, y = ed50_adj)) +
  geom_point(size = 0.75) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_few() +
  theme(axis.text = element_text(size = 7)) +
  scale_x_continuous(labels = function(x) str_trim(x)) +
  labs(x = "MMM (5 km)", y = NULL)

cowplot::plot_grid(mmm9.plot, mmm5.plot, ncol = 2)
```

```{r, fig.width = 11, fig.height = 3}
fig5 <- cowplot::plot_grid(lat.plot, lon.plot, mmm9.plot, mmm5.plot, nrow = 1,
                   rel_widths = c(1.07,1,1,1), labels = "AUTO")
ggsave(plot = fig5, filename = "output/fig5.png", width = 190, height = 50, units = "mm")
```

```{r}
mod <- lm(ed50_adj ~ source_lat * source_lon * mmm_9km * mmm_5km, data = dft)
anova(mod) %>%
  tidy() %>%
  knitr::kable()
```

There are no relationships between ED50 and source colony latitude, longitude, or MMM temperature. This is a pretty cool finding because it suggests that thermally tolerant colonies are not restricted to specific locations, but can be found throughout the Florida reef tract.

# Is the average ED50 of corals at each nursery related to the nursery's MMM temperature?
```{r}
# Relationship between nursery MMM and average nursery ED50?
nursmmm <- dft %>%
  group_by(nursery) %>%
  summarize(meaned50 = mean(estimate_ed50)) %>%
  mutate(nursmmm = case_when(nursery == "mote" ~ 30.2694,
                             nursery == "um"  ~ 29.5412,
                             nursery == "nsu" ~ 29.5344,
                             nursery == "crf" ~ 29.7307,
                             nursery == "rrt" ~ 29.7307,
                             nursery == "fwc" ~ 29.9965)) 

nursmmm %>%
  ggplot(aes(x = nursmmm, y = meaned50)) +
  geom_label(aes(label = nursery))
```


```{r, eval = FALSE, include = FALSE}
top10 <- res.adj %>%
  filter(ed50_adj > quantile(ed50_adj, 0.9))
bot10 <- res.adj %>%
  filter(ed50_adj < quantile(ed50_adj, 0.1))


ggplot(res.adj, aes(x = source_lon, y = source_lat)) + 
  geom_point(size = 0.25, color = "gray") +
  geom_point(data = top10, color = "red", size = 2, alpha = 0.5) +
  #geom_point(data = bot10, color = "royalblue", size = 2, alpha = 0.5) +
  theme_few()

ggmap(map_fl, extent = "device") +
  geom_point(aes(x = source_lon, y = source_lat), data = res.adj, size = 0.25, color = "gray") +
  geom_point(aes(x = source_lon, y = source_lat), data = top10, color = "red", size = 2, alpha = 0.5)

top10

library(maps)
library(sf) 
library(units)
library(rnaturalearth)
dd <- st_as_sf(top10, coords = c("source_lon", "source_lat"), crs = 4326)
st_distance(dd)

top10 %>%
  count(nursery)
bot10 %>%
  count(nursery)

res.adj %>% anti_join(top20) %>% anti_join(bot20) %>%
  count(nursery)

res.adj %>%
  filter(nursery == "fwc") %>%
  inner_join(top20)
```


# Compare October to June CBASS results

27 of the genotypes we tested on the October cruise were also tested in June 2020 by Kelsey Johnson-Sapp at RSMAS in an independent set of CBASS assays (this occurred at the time of the genotype swap). We can compare the ED50 values obtained for these 27 genotypes in June and October to see how reproducible the results are.

```{r}
# Filter data for corals run in October and June
df2 <- df.all %>%
  drop_na(date) %>%
  group_by(nursery, geno) %>%
  filter(n_distinct(date) == 2) %>%
  ungroup() %>%
  select(nursery, name, geno, date, max_temp, fvfm)

# Fit dose-response curves for each genotype on each date
df2$genoN <- factor(paste0("g", group_indices(df2, nursery, geno, date)))
mod2 <- drm(fvfm ~ max_temp, curveid = genoN, data = df2, 
           fct = LL.3(names = c("hill", "max", "ED50")),
           upperl = c(120, 0.72, 40),
           lowerl = c(10, 0.55, 30))

# Extract ED50 estimates and standard error
ed50s <- tidy(mod2) %>%
  filter(term == "ED50") %>%
  select(genoN = curve, estimate, std.error) %>%
  left_join(distinct(df2, nursery, geno, date, genoN))

# Rearrange data to compare ED50 values between dates
ed50s2 <- ed50s %>% 
  mutate(date = recode(date, "2020-10" = "oct", "2020-06" = "jun")) %>%
  select(nursery, geno, date, ed50 = estimate, std.error) %>%
  pivot_wider(names_from = date, values_from = c(ed50, std.error)) %>%
  filter(geno != "FM19")

# Plot figure
fig <- ggplot(ed50s2, aes(y = ed50_oct, x = ed50_jun)) +
  geom_errorbar(aes(xmin = ed50_jun - std.error_jun,
                    xmax = ed50_jun + std.error_jun), lwd = 0.15) +
  geom_errorbar(aes(ymin = ed50_oct - std.error_oct,
                    ymax = ed50_oct + std.error_oct), lwd = 0.15) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  #geom_label(aes(label = geno), alpha = 0.7, size = 2) +
  annotate(geom = "text", x = 34.8, y = 36.5, 
           label = "Pearson's correlation:", adj = 0, size = 2) +
  stat_cor(method = "pearson", label.y = 36.45, label.x = 34.85, size = 2) +
  annotate(geom = "text", x = 34.8, y = 36.375, 
           label = "Spearman's rank correlation:", adj = 0, size = 2) +
  stat_cor(method = "spearman", label.y = 36.325, label.x = 34.85, size = 2) +
  geom_abline(intercept = 0, slope = 1, lty = 2, color = "black") +
  coord_fixed()  +
  labs(x = "ED50 (°C) measured in June", y = "ED50 (°C) measured in Oct.") +
  theme_few()
fig

ggsave("output/fig7.png", width = 90, height = 99, units = "mm")

# Linear regression
mod <- lm(ed50_oct ~ ed50_jun, data = ed50s2)
## Slope of fit
tibble(slope = coef(mod)[2]) %>% knitr::kable()
## Test if slope is different than 1
car::linearHypothesis(mod, "ed50_jun = 1")
## Find mean difference between Oct and Jun values
augment(mod) %>%
  mutate(diff = ed50_oct - ed50_jun) %>%
  summarize("Mean difference Oct. - Jun" = mean(diff)) %>%
  knitr::kable()
```

There is a high correlation between ED50 values obtained from the same coral genets (from the same nurseries) run in both June and October (n = 27 genets). This suggests that ED50 values from independently run CBASS assays are relatively reproducible for a given genotype. The shift in y-intercept could be due to seasonal differences, or experimental differences, in October vs. June. Overall, this high reproducibility is very encouraging for the continued use of the CBASS approach for thermal tolerance phenotyping.

# Thermal tolerance phenotype rarefaction curves
```{r}
# ED50 rarefaction curves from each nursery
adjed50dist <- rnorm(100000, meanadj, sdadj)
deciles <- quantile(adjed50dist, probs = seq(0, 1, 0.1))

sample_ed50s <- function(ed50s, n) {
  s <- sample(ed50s, n)
  prop_deciles <- length(unique(cut(s, deciles))) / 10
  return(prop_deciles)
}

rc <- res.adj %>%
  select(nursery, ed50_adj) %>%
  nest(data = ed50_adj) %>%
  crossing(n = 1:25) %>%
  mutate(prop = map2(data, n, ~ replicate(1000, sample_ed50s(.x$ed50_adj, .y))),
         mean = map_dbl(prop, mean))

rc1fig <- ggplot(rc, aes(x = n, y = mean, color = nursery)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(limits = c(0, 1)) +
  theme_few() +
  theme(axis.text = element_text(size = 7),
        legend.position = "none") +
  labs(x = "Genets sampled", y = "Proportion of ED50 range\nin sample")



# Prob of any genos in top 25% of population
sample_ed50s2 <- function(ed50s, n) {
  top25 <- quantile(rnorm(100000, mean(ed50s), sd(ed50s)), prob = 0.75)
  s <- sample(ed50s, n)
  any_top25 <- any(s >= top25)
  return(any_top25)
}

rc2 <- res.adj %>%
  #mutate(ed50_adj = rnorm(n = 229, mean = 35.98, sd = 10)) %>%
  select(nursery, ed50_adj) %>%
  nest(data = ed50_adj) %>%
  crossing(n = 1:25) %>%
  mutate(ntop25 = map2(data, n, ~ replicate(100, sample_ed50s2(.x$ed50_adj, .y))),
         mean = map_dbl(ntop25, mean))

rc2fig <- ggplot(rc2, aes(x = n, y = mean, color = nursery)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(limits = c(0, 1)) +
  theme_few() +
  theme(axis.text = element_text(size = 7),
        legend.position = c(0.8, 0.4)) +
  labs(x = "Genets sampled", y = "Probability sample includes\ntop quartile ED50")

fig7 <- cowplot::plot_grid(rc1fig, rc2fig, ncol = 2, labels = "AUTO")
ggsave("output/fig6.png", width = 172, height = 75, units = "mm")
```


```{r, include = FALSE, eval = FALSE}
myd <- rnorm(n = 1000000, mean = 10, sd = 0.1)
myq <- quantile(rnorm(1000000, mean(myd), sd(myd)), prob = 0.75)
myq

sample_ed50s2 <- function(ed50s, n) {
  s <- sample(ed50s, n)
  any_top25 <- any(s >= myq)
  return(any_top25)
}


tibble(n = 1:20) %>%
  mutate(ntop25 = map(n, ~ replicate(1000, sample_ed50s2(myd, .x))),
         mean = map_dbl(ntop25, mean))

test <- c("a", "b", "c", "d")
draw7 <- replicate(100000, any(sample(test, 8, replace = TRUE) == "d"))
table(draw7)
```

```{r}
res %>%
  group_by(nursery) %>%
  summarize(var = var(estimate_ed50))

var(c(35, 37))
var(c(35, 36, 37, 38))

var(c(35, 36))
var(c(36, 37))
var(c(35, 37))
var(c(35, 38))
var(c(36, 38))
var(c(37, 38))


myd <- rnorm(n = 40, mean = 35.98, sd = 0.38)

pvar <- function(ed50s, n) {
  s <- sample(ed50s, n)
  svar <- var(s)
  tvar <- var(ed50s)
  return(svar / tvar)
}


tibble(n = 2:20) %>%
  mutate(hi = map(n, ~ replicate(100, pvar(myd, .x))),
         mean = map_dbl(hi, mean))
```

```{r}
# ED50 rarefaction curves from each nursery

sample_ed50s <- function(ed50s, n) {
  s <- sample(ed50s, n)
  rs <- diff(range(s))
  #rt <- diff(range(ed50s))
  rt <- diff(qnorm(c(0.01, 0.99), mean(ed50s), sd(ed50s)))
  propr <- rs / rt
  return(propr)
}

rc <- res %>%
  select(nursery, estimate_ed50) %>%
  nest(data = estimate_ed50) %>%
  crossing(n = 1:25) %>%
  mutate(prop = map2(data, n, ~ replicate(10000, sample_ed50s(.x$estimate_ed50, .y))),
         mean = map_dbl(prop, mean))

rc1fig <- ggplot(rc, aes(x = n, y = mean, color = nursery)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(limits = c(0, 1)) +
  theme_few() +
  theme(axis.text = element_text(size = 7)) +
  labs(x = "Genets sampled", y = "Proportion of each nursery's\n total ED50 range captured in sample")
rc1fig
```


```{r}
# standard deviation of sample of X genets?

sample_ed50s <- function(ed50s, n) {
  m <- mean(ed50s)
  s <- sample(ed50s, n)
  sds <- sum(sapply(s, function(s) (s - m)^2))
  sda <- sd(ed50s)
  return(sds / sda)
}

rc <- res %>%
  select(nursery, estimate_ed50) %>%
  nest(data = estimate_ed50) %>%
  crossing(n = 1:25) %>%
  mutate(prop = map2(data, n, ~ replicate(100, sample_ed50s(.x$estimate_ed50, .y))),
         mean = map_dbl(prop, mean))

rc1fig <- ggplot(rc, aes(x = n, y = mean, color = nursery)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(limits = c(0, 1)) +
  theme_few() +
  theme(axis.text = element_text(size = 7)) +
  labs(x = "Genets sampled", y = "Proportion of each nursery's\n total ED50 range captured in sample")
rc1fig
```